### 2022-12-15 Rmd/runDD.Rmd
1st push

---
title: "**Density Dependence**"
subtitle: "Simulations"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
mathjax: TRUE

tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
# bibliography: refs.bib
---

This document provides an example of how to model density dependence in mass-at-age and M-at-age.

The life history parameters of mackerel are used to construct an `FLBRP` object that represents the equilibrium dynamics, this is then used to construct an `FLStock` object that models the time series dynamics.

The Lorenzen relationship between mass-at-age and M is used to link M and weight-at-age. For example, lorenzen M = a*300^-0.288, so if you want M=0.2 at 300g then a= 0.2 /(300^-0.288)

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/dd/",
               fig.path  ="../tex/figs/dd/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```

```{r, pkgs}
library(FLCore)
library(FLBRP)
library(FLasher)
library(FLife)
library(ggplotFL)

library(popbio)

library(GGally)
library(ggpubr)

library(rjson) # install.packages("rjson"); used for dropbox function
library(RJSONIO)

library(plyr)
library(dplyr)
library(reshape)

library(readxl)

library(mgcv)
library(mgcViz)

theme_set(theme_bw(16))
```

```{r, dbox}
dropboxdir<-try(file.path(get_dropbox(), "dd"))

if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

# Mackerel

```{r, mac.data}
load(file.path(dropboxdir,"data/om/mac.RData"))
ts =as.data.frame(read_excel("../data/inputs/dat.xlsx",3))
```

```{r eq}
par=teleost[c("linf","k","t0","l50","a","b"),dimnames(teleost)$iter=="Scomber scombrus"]
par=lhPar(par,m1=0.2/(300^-0.288),m2=-0.28,m3=NA)
eq =lhEql(par,m="lorenzen")

plot(eq,refpt="msy")
```

**Figure `r iFig=iFig+1; iFig`** Equilibrium dynamics

```{r}
om=as(eq,"FLStock")
om=fwd(om,fbar=fbar(om)[,-1],sr=eq)

plot(om)
```

**Figure `r iFig=iFig+1; iFig`** Time series dynamics

```{r}
plot(qmax((biomass(eq)%/%biomass(eq)[,c(14)])^0.3,0.5))+
  scale_y_continuous(labels = scales::percent)+
  xlab("Biomass")+ylab("% Change in mass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Density dependence multiplier

```{r}
ddWt<-function(wt,biomass, bmsy)
  wt%*%qmax((biomass/bmsy)^0.3,0.5)

wt=ddWt(stock.wt(eq),biomass(eq),refpts(eq)["msy","biomass"])

ggplot(as.data.frame(wt))+           
    geom_line(aes(age,data,col=factor(year)))+
    theme(legend.position="none")+
    scale_x_continuous(limits=c(0,15))+
    xlab("Age")+ylab("NMass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Simulation of density dependence in mass-at-age

```{r}
m=par["m1"]%*%(wt%^%par["m2"])

ggplot(as.data.frame(m))+           
    geom_line(aes(age,data,col=factor(year)))+
    theme(legend.position="none")+
    scale_x_continuous(limits=c(0,15))+
    xlab("Age")+ylab("M-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Simulation of density dependence in M-at-age

```{r}
x=propagate(eq,length(dimnames(eq)$year))
fbar(x)=fbar(x)[,1]
fbar(x)=c(fbar(eq))

for (i in seq(20)){
stock.wt(x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"])
m(       x)=par["m1"]%*%(stock.wt(x)%^%par["m2"])
x=brp(x)}

ggplot()+
  geom_line(aes(biomass,catch),
            data=model.frame(FLQuants(x,biomass=function(x) biomass(x) ,catch=function(x) catch(x)),drop=T))+
  geom_line(aes(biomass,catch),
            data=model.frame(FLQuants(eq,biomass=function(x) biomass(x) ,catch=function(x) catch(x)),drop=T),col="red")
```

**Figure `r iFig=iFig+1; iFig`** Comparison of equilibrium curves with (black) and without (red) density dependence.

```{r}
ggplot()+
  geom_line(aes(ssb,catch),
            data=model.frame(FLQuants(x,ssb=function(x) ssb(x) ,catch=function(x) catch(x)),drop=T))+
  geom_line(aes(ssb,catch),
            data=model.frame(FLQuants(eq,ssb=function(x) ssb(x) ,catch=function(x) catch(x)),drop=T),col="red")
```

**Figure `r iFig=iFig+1; iFig`** Comparison of equilibrium curves with (black) and without (red) density dependence.

# Funding

# References