### 2022-12-15 Rmd/runCond.Rmd
1st

---
title: "**Density Dependence**"
subtitle: "Summary"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE

tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
# bibliography: refs.bib
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)
opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/cond/",
               fig.path  ="../tex/figs/cond/",
               fig.width =8,
               fig.height=6,
               dev       ="png")
iFig=0
iTab=0
```

```{r, pkgs}
library(FLCore)
library(ggplotFL)
library(GGally)
library(ggpubr)
library(rjson) # install.packages("rjson"); used for dropbox function
library(RJSONIO)
library(plyr)
library(dplyr)
library(reshape)
library(readxl)

library(mgcv)
library(mgcViz)

theme_set(theme_bw(16))
```

```{r, eval=FALSE}
source("../R/get_dropbox.R")
```

```{r, dbox}
dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

# Mackerel

```{r, mac.data}
load(file.path(dropboxdir,"data/om/mac.RData"))
ts =as.data.frame(read_excel(file.path(dropboxdir,"data/inputs/dat.xlsx"),3))
wt =transform(as.data.frame(stock.wt(mac),drop=TRUE),cohort=year-as.numeric(age))
wt =merge(wt,ts,by="year")
wt=ddply(wt,.(age), transform, rsdl =data/mean(data))
wt=ddply(wt,.(age), transform, rsdl_=rsdl/var(data)^0.5)
wt$age=factor(wt$age,levels=sort(unique(wt$age)))
mat=transform(as.data.frame(mat(mac),drop=TRUE),cohort=year-as.numeric(age))
mat=merge(mat,ts)
mat=ddply(mat,.(age), transform, rsdl =data/mean(data))
mat=ddply(mat,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
mat$age=factor(mat$age,levels=sort(unique(mat$age)))
m  =transform(as.data.frame(m(mac),drop=TRUE),cohort=year-as.numeric(age))  
m  =merge(m,ts)
m  =ddply(m,.(age), transform, rsdl =data/mean(data))
m  =ddply(m,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
m$age=factor(m$age,levels=sort(unique(m$age))) 
```

```{r, mac.ts, fig.width=8,fig.height=4}
ggplot(transform(melt(ts,id="year"),year=year,quant=variable,data=value))+
  geom_line(aes(year,data,col=quant))+
  theme(legend.position="bottom")+
  scale_color_manual("Quantity",labels=c("SSB","Total Biomass","ICES Biomass"),values=rainbow(3))+
  xlab("Year")+ylab("")
```

**Figure `r iFig=iFig+1; iFig`** Time series

```{r, mac.wt, fig.width=8,fig.height=4}
ggplot(wt)+  
  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+
  geom_line(aes(year,data,group=age,col=age))+
  geom_point(aes(year,data,col=age))+
  xlab("Year")+ylab("Mass-at-age")+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age

```{r, mac.m, fig.width=8,fig.height=4}
ggplot(mat)+  
  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+ 
  geom_line(aes(year,data,group=age,col=age))+
  geom_point(aes(year,data,col=age))+
  xlab("Year")+ylab("Mass-at-age")+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age

```{r, mac.mat, fig.width=8,fig.height=4}
ggplot(m)+  
  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+ 
  geom_line(aes(year,data,group=age,col=age))+
  geom_point(aes(year,data,col=age))+
  xlab("Year")+ylab("Mass-at-age")+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
```

**Figure `r iFig=iFig+1; iFig`** M-at-age

```{r, mac.wt2, fig.width=8,fig.height=5}
#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
a=ggplot(wt)+
  geom_boxplot(aes(age,data))+
  xlab("Mass-at-age")+ylab("Age")
b=ggplot(subset(wt,rsdl<3&rsdl>0))+
  geom_point(aes(tb,rsdl))+
  geom_smooth(aes(tb,rsdl),method="lm")+
  facet_wrap(age~.)+
  xlab("Biomass")+ylab("Residual")
ggarrange(a,b ,label.y=1,  
          widths = c(4,6), heights = c(1,1,1),
          nrow   = 1,          ncol    = 2)+
  theme_bw(16)+
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age, summary

```{r, mac.wt.tb, fig.height=8,fig.width=8}
ggplot(subset(wt,age!=0),aes(tb,data))+
  geom_point()+
  geom_smooth(se=F,method="lm")+
  facet_wrap(~age,scale="free",ncol=3)+
  xlab("Biomass")+ylab("Mass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass

```{r, mac.mat2, fig.width=8,fig.height=5}
#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
a=ggplot(mat)+
  geom_boxplot(aes(age,data))+
  xlab("Mass-at-age")+ylab("Age")
b=ggplot(mat)+
  geom_point(aes(tb,rsdl))+
  geom_smooth(aes(tb,rsdl),method="lm")+
  facet_wrap(age~.)+
  xlab("Biomass")+ylab("Residual")
ggarrange(a,b ,label.y=1,  
          widths = c(4,6), heights = c(1,1,1),
          nrow   = 1,          ncol    = 2)+
  theme_bw(16)+
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age, summary

```{r, mac.mat.tb, fig.height=5,fig.width=8}
ggplot(subset(mat,age%in%(1:5)),aes(tb,data))+
  geom_point()+
  geom_smooth(se=F,method="lm")+
  facet_wrap(~age,scale="free",ncol=3)+
  xlab("Biomass")+ylab("Maturity-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age v Biomass

```{r}
gmr=gammV(data~s(tb, bs="tp")+age, data=wt)
summary(gmr)
plot(gmr)
```

# Blue whiting

```{r, whb.data}
load(file.path(dropboxdir,"data/om/whb.RData"))
ts=model.frame(FLQuants(whb,tb=biomass,ssb=ssb),drop=T)
ts$tb=ts$tb/1e6
wt =transform(as.data.frame(stock.wt(whb),drop=TRUE),cohort=year-as.numeric(age))
wt =merge(wt,ts,by="year")
wt=ddply(wt,.(age), transform, rsdl =data/mean(data))
wt=ddply(wt,.(age), transform, rsdl_=rsdl/var(data)^0.5)
wt$age=factor(wt$age,levels=sort(unique(wt$age)))
mat=transform(as.data.frame(mat(whb),drop=TRUE),cohort=year-as.numeric(age))
mat=merge(mat,ts)
mat=ddply(mat,.(age), transform, rsdl =data/mean(data))
mat=ddply(mat,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
mat$age=factor(mat$age,levels=sort(unique(mat$age)))
m  =transform(as.data.frame(m(whb),drop=TRUE),cohort=year-as.numeric(age))  
m  =merge(m,ts)
m  =ddply(m,.(age), transform, rsdl =data/mean(data))
m  =ddply(m,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
m$age=factor(m$age,levels=sort(unique(m$age)))
```

```{r, whb.ts, fig.width=8,fig.height=4}
ggplot(transform(melt(ts,id="year"),year=year,quant=variable,data=value))+
  geom_line(aes(year,data,col=quant))+
  theme(legend.position="bottom")+
  scale_color_manual("Quantity",labels=c("SSB","Total Biomass","ICES Biomass"),values=rainbow(3))+
  xlab("Year")+ylab("")
```

**Figure `r iFig=iFig+1; iFig`** Time series

```{r, whb.wt, fig.width=8,fig.height=4}
ggplot(wt)+  
  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+
  geom_line(aes(year,data,group=age,col=age))+
  geom_point(aes(year,data,col=age))+
  xlab("Year")+ylab("Mass-at-age")+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age

```{r, whb.mat, fig.width=8,fig.height=4}
ggplot(mat)+  
  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+
  geom_line(aes(year,data,group=age,col=age))+
  geom_point(aes(year,data,col=age))+
  xlab("Year")+ylab("Mass-at-age")+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age

```{r, whb.m, fig.width=8,fig.height=4}
ggplot(m)+  
  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+ 
  geom_line(aes(year,data,group=age,col=age))+
  geom_point(aes(year,data,col=age))+
  xlab("Year")+ylab("Mass-at-age")+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
```

**Figure `r iFig=iFig+1; iFig`** M-at-age

```{r, whb.wt2, fig.width=8,fig.height=5}
#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
a=ggplot(wt)+
  geom_boxplot(aes(age,data))+
  xlab("Mass-at-age")+ylab("Age")
b=ggplot(subset(wt,rsdl<3&rsdl>0))+
  geom_point(aes(tb,rsdl))+
  geom_smooth(aes(tb,rsdl),method="lm")+
  facet_wrap(age~.,ncol=3)+
  xlab("Biomass")+ylab("Residual")
ggarrange(a,b ,label.y=1,  
          widths = c(4,6), heights = c(1,1,1),
          nrow   = 1,          ncol    = 2)+
  theme_bw(16)+
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age, summary

```{r, whb.wt.tb, fig.height=8,fig.width=8}
ggplot(wt,aes(tb,data))+
  geom_point()+
  geom_smooth(se=F,method="lm")+
  facet_wrap(~age,scale="free",ncol=3)+
  xlab("Biomass")+ylab("Mass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass

```{r, whb.mat2, fig.width=8,fig.height=5}
#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
a=ggplot(mat)+
  geom_boxplot(aes(age,data))+
  xlab("Mass-at-age")+ylab("Age")
b=ggplot(mat)+
  geom_point(aes(tb,rsdl))+
  geom_smooth(aes(tb,rsdl),method="lm")+
  facet_wrap(age~.,ncol=3)+
  xlab("Biomass")+ylab("Residual")
ggarrange(a,b ,label.y=1,  
          widths = c(4,6), heights = c(1,1,1),
          nrow   = 1,          ncol    = 2)+
  theme_bw(16)+
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age, summary

```{r, whb.mat.tb, fig.height=5,fig.width=8}
ggplot(subset(mat,age%in%1:5),aes(tb,data))+
  geom_point()+
  geom_smooth(se=F,method="lm")+
  facet_wrap(~age,scale="free",ncol=3)+
  xlab("Biomass")+ylab("Mass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass

```{r}
gmr=gammV(data~s(tb)+age, data=wt)
summary(gmr)
plot(gmr)
```

# Funding

# References