### 2023-01-10 Rmd/mseWhb.Rmd
cleaned up

---
title: "**Density Dependence**"
subtitle: "Example of MSE for Blue whiting"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
bibliography: refs.bib
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/mseWhb/",
               fig.path  ="../tex/figs/mseWhb/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```

```{r, pkgs}
library(FLCore)
library(FLBRP)
library(FLasher)
library(FLAssess)
library(FLife)
library(ggplotFL)

library(popbio)

library(GGally)
library(ggpubr)

library(rjson) # install.packages("rjson"); used for dropbox function
library(RJSONIO)

library(plyr)
library(dplyr)
library(reshape)

library(readxl)

library(mgcv)
library(mgcViz)

theme_set(theme_bw(16))
```

```{r}
source("../R/dd.R")
```

```{r, dbox}
dropboxdir<-try(file.path(get_dropbox(), "dd"))

if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

```{r, whb.data}
load(file.path(dropboxdir,"data/om/whb.RData"))

par=FLPar(c(m1=0.08/(0.2^-0.288),m2=-0.28,bmsy=6e6,k=63,w50=0.08))

stock.wt(   whb)[1]=0.01
catch.wt(   whb)[1]=0.01
landings.wt(whb)[1]=0.01
discards.wt(whb)[1]=0.01

om=ddFn2(whb,par)
om=om+VPA(om)
om=ddFn2(om,par)

sr=fmle(as.FLSR(om,model="bevholt"),control=list(silent=TRUE))
eq=FLBRP(om,sr=sr)

om=fwdWindow(om,end=2080,eq)
f =fbar(om)[,ac(2021:2080)]%=%refpts(eq)["msy","harvest"]
om=fwd(om,fbar=f,sr=eq)

om2=om
om3=om
om4=om
for (iYear in ac(2021:2080)){
  om2=ddFn(iYear,om2,par)
  om2=fwd(om2,fbar=f[,iYear]*0.8,sr=eq)

  om3=ddFn(iYear,om3,par)
  om3=fwd(om3,fbar=f[,iYear],sr=eq)

  om4=ddFn(iYear,om4,par)
  om4=fwd(om4,fbar=f[,iYear]*1.2,sr=eq)
  }

plot(FLStocks("fwd"=om,"0.8"=om2,"1.0"=om3,"1.2"=om4))
```

```{r}
ggplot(FLQuants("ICES"=stock.wt(om[ ,"2080"]), 
                "0.8" =stock.wt(om2[,"2080"]), 
                "1.0" =stock.wt(om3[,"2080"]),
                "1.2" =stock.wt(om4[,"2080"])))+
  geom_line(aes(age,data,col=qname))+
  xlab("Age")+ylab("Mass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age

```{r}
ggplot(FLQuants("ICES"=m(om[ ,"2080"]),  
                "0.8" =m(om2[,"2080"]), 
                "1.0" =m(om3[,"2080"]),
                "1.2" =m(om4[,"2080"])))+
  geom_line(aes(age,data,col=qname))+
  xlab("Age")+ylab("M") 
```

**Figure `r iFig=iFig+1; iFig`** M-at-mass

```{r}
ggplot(FLQuants("ICES"=mat(om[ ,"2080"]),  
               "0.8"  =mat(om2[,"2080"]),
               "1.0"  =mat(om3[,"2080"]), 
               "1.2"  =mat(om4[,"2080"])))+
  geom_line(aes(age,data,col=qname))+
  xlab("Age")+ylab("Maturity-a-age")
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age

```{r}
ggplot(rbind(cbind(What="ICES",model.frame(FLQuants(Mat=mat(om[  ,"2080"]),Mass=stock.wt(om[ ,"2080"])))),  
             cbind(What= "0.8",model.frame(FLQuants(Mat=mat(om2[ ,"2080"]),Mass=stock.wt(om2[,"2080"])))),
             cbind(What= "1.0",model.frame(FLQuants(Mat=mat(om3[ ,"2080"]),Mass=stock.wt(om3[,"2080"])))),
             cbind(What= "1.2",model.frame(FLQuants(Mat=mat(om4[ ,"2080"]),Mass=stock.wt(om4[,"2080"]))))))+
  geom_line(aes(Mass,Mat,col=What))+
  geom_point(aes(Mass,Mat,col=What))
```

**Figure `r iFig=iFig+1; iFig`** M-at-mass