### 2023-02-22 Rmd/mseMac.Rmd
with alpha set by data

---
title: "**Density Dependence**"
subtitle: "MSE for Mackerel"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
bibliography: refs.bib
---

## Conditioning procedure

+  Load ICES assessment
+  Time varying M
+  Conduct VPA

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/mseMac/",
               fig.path  ="../tex/figs/mseMac/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```

```{r, pkgs}
library(FLCore)    # install.packages("FLCore", repos="http://flr-project.org/R")
library(FLBRP)     # install.packages("FLBRP", repos="http://flr-project.org/R")
library(FLasher)   # install.packages("FLasher", repos="http://flr-project.org/R")
library(FLAssess)  # install.packages("FLAssess", repos="http://flr-project.org/R")
library(FLife)     # install.packages("FLife", repos="http://flr-project.org/R")
library(ggplotFL)

library(popbio)

library(GGally)
library(ggpubr)

library(rjson) # install.packages("rjson"); used for dropbox function
library(RJSONIO)

library(plyr)
library(dplyr)
library(reshape)

library(readxl)

library(mgcv)
library(mgcViz)

theme_set(theme_bw(16))
```

```{r}
source("../R/dd.R")
source("../R/get_dropbox.r")
```

```{r, dbox}
dropboxdir<-try(file.path(get_dropbox(), "pelagics"))

if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

```{r, ices}
stock="Mac"
load(paste("../data/om/ices",stock,".RData",sep=""))
stk=get(tolower(stock))

stock.wt(   stk)[1]=0.01
catch.wt(   stk)[1]=0.01
landings.wt(stk)[1]=0.01
discards.wt(stk)[1]=0.01
```

```{r, pars}
par=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.288,bref=NA,b=-0.3,matk=30.1,w50=0.2))
```

```{r, m}
m(stk)=par["m1"]%*%(stock.wt(stk)%^%par["m2"])
vpa=stk+VPA(stk)
```

```{r, vpa}
plot(FLStocks("ICES"=stk,"VPA"=vpa))
```

```{r, sr}
load(paste("/home/laurie/Dropbox/pelagics/results/mac/figures/",tolower(stock),"_objects.RData",sep=""))

stk=window(iter(vpa,1),start=2000,end=2019)
sr=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
                control=list(silent=TRUE))
eq=FLBRP(stk,sr=ab(sr))

stk=window(iter(vpaM,1),start=2000,end=2019)
srM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
                control=list(silent=TRUE))
eqM=FLBRP(stk,sr=ab(srM))

stk=window(iter(vpaDDM,1),start=2000,end=2019)
srDDM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
                control=list(silent=TRUE))
eqDDM=FLBRP(stk,sr=ab(srDDM))

stk=window(iter(vpaDDMM,1),start=2000,end=2019)
srDDMM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
                control=list(silent=TRUE))
eqDDMM=FLBRP(stk,sr=ab(srDDMM))

stk=window(iter(vpaDDMMM,1),start=2000,end=2019)
srDDMMM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
                control=list(silent=TRUE))
eqDDMMM=FLBRP(stk,sr=ab(srDDMMM))

dat=rbind(
  cbind(What="ICES",    model.frame(FLQuants(eq,rec=rec,ssb=ssb,catch=catch),drop=T)),
  cbind(What="VPA",     model.frame(FLQuants(eqM,rec=rec,ssb=ssb,catch=catch),drop=T)),
  cbind(What="VPA M",   model.frame(FLQuants(eqDDM,rec=rec,ssb=ssb,catch=catch),drop=T)),
  cbind(What="VPA MM",  model.frame(FLQuants(eqDDMM,rec=rec,ssb=ssb,catch=catch),drop=T)),
  cbind(What="VPA MMM", model.frame(FLQuants(eqDDMMM,rec=rec,ssb=ssb,catch=catch),drop=T)))
```

```{r, eqs}
ggplot(dat)+
  geom_line(aes(ssb,catch,col=What))
```

**Figure `r iFig=iFig+1; iFig`** Stock recruitment relationships

```{r, srs}
ggplot(dat)+
  geom_line(aes(ssb,rec,col=What))
```

**Figure `r iFig=iFig+1; iFig`** Equilibrium curves for assumptions about density dependence

```{r, ddM}
par["bref"]=refpts(eqM)["msy","biomass"]

for (i in dimnames(vpaM)$year)
  om=ddFn(i,vpaM,par,FALSE,FALSE,TRUE)
```

```{r, omCheck}
om=fwdWindow(window(om,end=2020),end=2080,eq)
f =FLQuant(rep(seq(0,c(refpts(eq)["crash","harvest"])*1.5,length.out=100),each=61),dimnames=dimnames(fbar(om)[,ac(2020:2080)]))
om=fwd(om,fbar=f,sr=eq)

oms=FLStocks("Base"     =om)
oms[["DD Mass"]]        =om
oms[["DD Mass, Mat"]]   =om
oms[["DD Mass, Mat, M"]]=om
for (iYear in ac(2020:2080)){
  oms[["DD Mass"]]=ddFn(iYear,oms[["DD Mass"]],par,TRUE,FALSE,FALSE)
  oms[["DD Mass"]]=fwd(oms[["DD Mass"]],fbar=f[,iYear],sr=eq)

  oms[["DD Mass, Mat"]]=ddFn(iYear,oms[["DD Mass, Mat"]],par,TRUE,TRUE,FALSE)
  oms[["DD Mass, Mat"]]=fwd(oms[["DD Mass, Mat"]],fbar=f[,iYear],sr=eq)

  oms[["DD Mass, Mat, M"]]=ddFn(iYear,oms[["DD Mass, Mat, M"]],par,TRUE,TRUE,TRUE)
  oms[["DD Mass, Mat, M"]]=fwd(oms[["DD Mass, Mat, M"]],fbar=f[,iYear],sr=eq)
  }
```

```{r, check2}
p1=ggplot(stock.wt(oms[[2]][,ac(2080)]))+ 
  geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(stock.wt(oms[[1]][,ac("2080")])),col="red")+
  xlab("Age")+ylab("Mass-at-age") 
p2=ggplot(mat(oms[[2]][,ac(2080)]))+
  geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(mat(oms[[1]][,ac("2080")])),col="red")+
  xlab("Age")+ylab("Mat-at-age")
p3=ggplot(m(oms[[2]][,ac(2080)]))+
  geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(m(oms[[1]][,ac("2080")])),col="red")+
  xlab("Age")+ylab("M-at-age")

d1=model.frame(FLQuants(oms[[1]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
d2=model.frame(FLQuants(oms[[2]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
p4=ggplot(d2)+
  geom_line(aes(SSB,Yield))+
  geom_line(aes(SSB,Yield),col="red",data=d1)

ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
```

**Figure `r iFig=iFig+1; iFig`** Equilibrium curves and biological parameters for assumptions about density dependence

```{r, check3}
p1=ggplot(stock.wt(oms[[3]][,ac(2080)]))+ 
  geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(stock.wt(oms[[1]][,ac("2080")])),col="red")+
  xlab("Age")+ylab("Mass-at-age") 
p2=ggplot(mat(oms[[3]][,ac(2080)]))+
  geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(mat(oms[[1]][,ac("2080")])),col="red")+
  xlab("Age")+ylab("Mat-at-age")
p3=ggplot(m(oms[[3]][,ac(2080)]))+
  geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(m(oms[[1]][,ac("2080")])),col="red")+
  xlab("Age")+ylab("M-at-age")

d1=model.frame(FLQuants(oms[[1]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
d3=model.frame(FLQuants(oms[[3]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
p4=ggplot(d3)+
  geom_line(aes(SSB,Yield))+
  geom_line(aes(SSB,Yield),col="red",data=d1)

ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
```

**Figure `r iFig=iFig+1; iFig`** Equilibrium curves and biological parameters for assumptions about density dependence

```{r, check4}
p1=ggplot(stock.wt(oms[[4]][,ac(2080)]))+ 
  geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(stock.wt(oms[[1]][,ac("2080")])),col="red")+
  xlab("Age")+ylab("Mass-at-age") 
p2=ggplot(mat(oms[[4]][,ac(2080)]))+
  geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(mat(oms[[1]][,ac("2080")])),col="red")+
  xlab("Age")+ylab("Mat-at-age")
p3=ggplot(m(oms[[4]][,ac(2080)]))+
  geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(m(oms[[1]][,ac("2080")])),col="red")+
  xlab("Age")+ylab("M-at-age")

d1=model.frame(FLQuants(oms[[4]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
d4=model.frame(FLQuants(oms[[2]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
p4=ggplot(d4)+
  geom_line(aes(SSB,Yield))+
  geom_line(aes(SSB,Yield),col="red",data=d1)

ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
```

**Figure `r iFig=iFig+1; iFig`** Equilibrium curves and biological parameters for assumptions about density dependence

```{r, eqlCurves}
rfs=ldply(oms, function(x) { 
  model.frame(FLQuants(x[,"2080"],
             biomass=function(x) stock(x),
             ssb    =function(x) ssb(  x),
             catch  =function(x) catch(x),
             f      =function(x) fbar( x)),drop=TRUE)[,-1]})

ggplot(rfs)+
  geom_line(aes(ssb,catch,col=.id))+
  xlab("SSB")+ylab("Catch")+
  theme(legend.position="bottom")
```

**Figure `r iFig=iFig+1; iFig`** Equilibrium curves for assumptions about density dependence

**Table `r iTab=iTab+1; iTab`** Reference points

```{r, refptsTable}
rfpts=ddply(rfs,.(.id), with, {  
  flag=catch==max(catch)
  data.frame(Bmsy  =biomass[flag],
             SSBmsy=ssb[flag],
             MSY   =catch[flag],
             Fmsy  =f[flag],
             Virgin=max(ssb),
             B0    =max(biomass))})

rfpts
```

```{r, projectons-1}
prj=FLStocks("1"=iter(oms[[1]],1:4),
             "2"=iter(oms[[2]],1:4),
             "3"=iter(oms[[3]],1:4),
             "4"=iter(oms[[4]],1:4))

F=fbar(prj[[1]][,ac(2020:2080)])%=%rep(rfpts$Fmsy,each=61)

control=as(FLQuants("f"=F),"fwdControl")
prj[[1]] =fwd(prj[[1]],control=control,sr=eq)

for (i in ac(2020:2080)) {
    prj[[2]] =ddFn(i,prj[[2]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
    control=as(FLQuants("f"=F[,i]),"fwdControl")
    prj[[2]] =fwd(prj[[2]],control=control,sr=eq)

    prj[[3]] =ddFn(i,prj[[3]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
    control=as(FLQuants("f"=F[,i]),"fwdControl")
    prj[[3]] =fwd(prj[[3]],control=control,sr=eq)

    prj[[4]] =ddFn(i,prj[[4]],par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
    control=as(FLQuants("f"=F[,i]),"fwdControl")
    prj[[4]] =fwd(prj[[4]],control=control,sr=eq)
    }
```

```{r}
plot(FLStocks(prj[[1]]),iter=1:4) 
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming no density dependence

```{r}
names(prj)=names(oms)
plot(FLStocks(prj[[2]]),iter=1:4)
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass-at-age only

```{r}
plot(FLStocks(prj[[3]]),iter=1:4)
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass and maturity-at-age only

```{r}
plot(FLStocks(prj[[4]]),iter=1:4)
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass, maturity and natural mortality-at-age

```{r, eval=FALSE}
ldply(prj,ldply, function(x) model.frame(FLQuants(x,biomass=function(x) biomass(x)[,"2080"],
                                                    ssb    =function(x) ssb(    x)[,"2080"],
                                                    catch  =function(x) catch(  x)[,"2080"],
                                                    f      =function(x) fbar(   x)[,"2080"]),drop=T)) 
```

```{r, projectons-2}
devRec=rlnorm(100,rec(prj[[1]])[,ac(2020:2080),,,,1]%=%0,0.3)

base=propagate(iter(prj[[1]],1),100)

prjs=mlply(data.frame(F=rfpts$Fmsy), function(F) {

            ## FMSY estimate by OM ######################################
            F=FLQuant(F,dimnames=list(year=2020:2080))

            ## No DD ####################################################
            control=as(FLQuants("f"=F),"fwdControl")
            x1     =fwd(base,control=control,sr=eq,residuals=devRec)

            ## With DD ####################################################
            x2=x1
            x3=x1
            x4=x1
            for (iYr in ac(2020:2080)) {
              ## DD M ##################################################         
              x2 =ddFn(iYr,x2,par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
              x2 =fwd(x2,control=control,sr=eq,residuals=devRec)

              ## DD MM##################################################
              x3 =ddFn(iYr,x3,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
              x3 =fwd(x3,control=control,sr=eq,residuals=devRec)

              ## DD MMM ################################################
              x4 =ddFn(iYr,x4,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
              x4 =fwd(x4,control=control,sr=eq,residuals=devRec)
              }  ## year loop

            FLStocks(list("Base"=x1,"M"=x2,"MM"=x3,"MMM"=x4))        
    }) ## F loop

save(prjs,file=paste(dirRuns,"/",stock,"Prjs.RData",sep=""))
```

```{r}
plot(FLStocks(prjs[[1]])) 
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at Base $F_{MSY}$

```{r}
plot(FLStocks(prjs[[2]])) 
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, with  $F_{MSY}$ for density dependence in mass-at-age

```{r}
plot(FLStocks(prjs[[3]])) 
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$ for density dependence in mass, maturity-at-age

```{r}
plot(FLStocks(prjs[[4]])) 
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$ for density dependence in mass, maturity, natural mortality-at-age

```{r, MSE, eval=FALSE}
mse=hcrICES(fmsy,eql,rec(fmsy),
                      par,
                      start,   end,
                      interval,lag=lag,
                      err=err,
                      bndTac=c(0,Inf))
```

```{r, chk}
dt1=model.frame(FLQuants("Rec"=rec(oms[[1]][,"2050"]),"SSB"=ssb(oms[[1]][,"2050"])),drop=T)
dt2=model.frame(FLQuants("Rec"=rec(vpa[,"2050"]),"SSB"=ssb(vpa[,"2050"])),drop=T)

ggplot()+
  geom_line(aes(SSB,Rec),data=dt1)+
  geom_line(aes(SSB,Rec),data=dt2)
```