### 2023-03-01 Rmd/omMac.Rmd

---
title: "**Operating model**"
subtitle: "Mackerel: Operating Model with density dependence in mass, maturity and M-at-age"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
mathjax: TRUE

tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
# bibliography: refs.bib
---

```{r warning=FALSE, cache=FALSE, include=FALSE, knitr, eval=TRUE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               cache     =TRUE,
               cache.path="../cache/mac/om/",
               fig.path  ="../tex/figs/mac/om/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```

```{r, pkgs}
library(FLCore)
library(FLBRP)
library(FLasher)
library(FLife)
library(ggplotFL)

library(popbio)

library(GGally)
library(ggpubr)

library(rjson) # install.packages("rjson"); used for dropbox function
library(RJSONIO)

library(plyr)
library(dplyr)
library(reshape)

library(readxl)

library(mgcv)
library(mgcViz)

library(compositions)

theme_set(theme_bw(16))
```

```{r, dropbox}
source("../R/get_dropbox.r")

dropboxdir<-try(file.path(get_dropbox(), "pelagics"))

if ("try-error"%in% is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

```{r echo=FALSE, warning=FALSE, cache=FALSE, eval=TRUE}

# source("~/Desktop/projects/ddmse/R/dd.R")
source("../R/dd.R")
```

```{r echo=FALSE, warning=FALSE, cache=FALSE, eval=TRUE}
load(file.path(dropboxdir,"data/om/mac.RData"))
```

```{r echo=FALSE, warning=FALSE, cache=FALSE, eval=TRUE}
load(file.path(dropboxdir,"data/inputs/ices.RData"))
ices=ices[["mac.27.nea"]]
ices=window(ices,start=1991)

sr  =as.FLSR(ices,model=bevholtSV)
sr  =fmle(sr,fixed=list(s=0.8,spr0=mean(spr0(ices))),control=list(silent=TRUE))
eq  =FLBRP(ices,ab(sr))
ices=fwdWindow(ices,eq,end=2050)

F  =propagate(window(fbar(ices),end=2050),101)
F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
          seq(c(refpts(eq)["msy",  "harvest"]),c(refpts(eq)["crash","harvest"])*1.2,length.out=51)[-1]),each=dim(F)[2])

control=as(FLQuants("f"=F[,-1]),"fwdControl")
ices   =fwd(propagate(ices,101),control=control,sr=eq)

dices=model.frame(FLQuants(ices[,"2050"],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                         "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
ggplot(dices)+
  geom_line(aes(SSB,Yield))
```

**Figure `r iFig=iFig+1; iFig`** .

```{r, vpa}
# Start the clock
ptm <- proc.time()

load(file.path(dropboxdir,"data/inputs/ices.RData"))
ices=ices[["mac.27.nea"]]
ices=window(ices,start=1991)
vpa =ices+FLAssess:::VPA(ices)

sr  =as.FLSR(vpa,model=bevholtSV)
sr  =fmle(sr,fixed=list(s=0.8,spr0=mean(spr0(vpa))),control=list(silent=TRUE))
eq  =FLBRP(vpa,ab(sr))
vpa =fwdWindow(vpa,eq,end=2050)

F  =propagate(window(fbar(vpa),start=2020),101)
F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
          seq(c(refpts(eq)["msy",  "harvest"]),c(refpts(eq)["crash","harvest"])*1.2,length.out=51)[-1]),each=dim(F)[2])

control=as(FLQuants("f"=F),"fwdControl")
vpa    =fwd(vpa,control=control,sr=eq)

dvpa=model.frame(FLQuants(vpa[,"2050"],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                       "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
pvpa=ggplot(dvpa)+
  geom_line(aes(SSB,Yield))

# Stop the clock
proc.time() - ptm
```

**Figure `r iFig=iFig+1; iFig`** .

```{r, vpaM}
# Start the clock
ptm <- proc.time()

#load(file.path(dropboxdir, "data/om/parMac.RData"))
load("/home/laurie/Desktop/projects/ddmse/data/om/parMac.RData")
par["m1"]  =par["m1"]/10
par["w50"] =par["w50"]/1000
par["matk"]=30

load(file.path(dropboxdir,"data/inputs/ices.RData"))
vpaM=ices[["mac.27.nea"]]
vpaM=window(vpaM,start=1991)
m(vpaM)=par["m1"]%*%(stock.wt(vpaM)%^%par["m2"])
vpaM=vpaM+FLAssess:::VPA(vpaM)

sr  =as.FLSR(vpaM,model=bevholtSV)
sr  =fmle(sr,fixed=list(s=0.8,spr0=mean(spr0(vpa))),control=list(silent=TRUE))
eq  =FLBRP(vpaM,ab(sr))
vpaM=fwdWindow(vpaM,eq,end=2050)

# plot(sr)
# plot(vpaM)

F  =propagate(window(fbar(vpaM),start=2020),101)
F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
          seq(c(refpts(eq)["msy",  "harvest"]),c(refpts(eq)["crash","harvest"])*1.5,length.out=51)[-1]),each=dim(F)[2])

control=as(FLQuants("f"=F),"fwdControl")
vpaM  =fwd(vpaM,control=control,sr=eq)

d1=model.frame(FLQuants(vpaM[,"2050"],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                      "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
d=rbind(cbind("What"="ICES",dices),
        cbind("What"="VPA",dvpa),
        cbind("What"="M",d1))
ggplot(d)+
  geom_line(aes(SSB,Yield,col=What))

# Stop the clock
proc.time() - ptm
```

\newpage

```{r}

as.data.frame(par) %>% pander::pandoc.table(style="simple")

```

\newpage

```{r, vpaDDM}

# source("~/Desktop/projects/ddmse/R/dd.R")

# Start the clock
ptm <- proc.time()

##################################
par["bref"]=subset(d1,Yield==max(Yield))[,"Biomass"]
##################################

vpaDDM=vpaM

for (i in ac(2020:2050)) {
  vpaDDM =ddFn(i,vpaDDM,par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
  control=as(FLQuants("f"=F[,i]),"fwdControl")
  vpaDDM =fwd(vpaDDM,control=control,sr=eq)}

p1=ggplot(stock.wt(vpaDDM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(stock.wt(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("Mass-at-age")
p2=ggplot(mat(vpaDDM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(mat(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("Mat-at-age")
p3=ggplot(m(vpaDDM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(m(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("M-at-age")

d2=model.frame(FLQuants(vpaDDM[,"2050"],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                        "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
p4=ggplot(d2)+
  geom_line(aes(SSB,Yield))+
  geom_line(aes(SSB,Yield),col="red",data=d1)

ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)

# Stop the clock
proc.time() - ptm
```

\newpage

```{r, vpaDDMM}

# Start the clock
ptm <- proc.time()

vpaDDMM =vpaM

for (i in ac(2020:2050)) {
  vpaDDMM =ddFn(i,vpaDDMM,par,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
  control=as(FLQuants("f"=F[,i]),"fwdControl")
  vpaDDMM =fwd(vpaDDMM,control=control,sr=eq)}

p1=ggplot(stock.wt(vpaDDMM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(stock.wt(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("Mass-at-age")
p2=ggplot(mat(vpaDDMM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(mat(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("Mat-at-age")
p3=ggplot(m(vpaDDMM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(m(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("M-at-age")

d3=model.frame(FLQuants(vpaDDMM[,"2050"],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                         "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
p4=ggplot(d3)+
  geom_line(aes(SSB,Yield))+
  geom_line(aes(SSB,Yield),col="red",data=d1)

ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)

# Stop the clock
proc.time() - ptm
```

**Figure `r iFig=iFig+1; iFig`** .

\newpage

```{r, vpaDDMMM}

# Start the clock
ptm <- proc.time()

vpaDDMMM=vpaM

for (i in ac(2020:2050)) {
  vpaDDMMM =ddFn(i,vpaDDMMM,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
  control=as(FLQuants("f"=F[,i]),"fwdControl")
  vpaDDMMM =fwd(vpaDDMMM,control=control,sr=eq)}

p1=ggplot(stock.wt(vpaDDMMM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(stock.wt(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("Mass-at-age")
p2=ggplot(mat(vpaDDMMM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(mat(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("Mat-at-age")
p3=ggplot(m(vpaDDMMM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(m(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("M-at-age")

d4=model.frame(FLQuants(vpaDDMMM[,"2050"],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                          "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
p4=ggplot(d4)+
  geom_line(aes(SSB,Yield))+
  geom_line(aes(SSB,Yield),col="red",data=d1)

ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)

# Stop the clock
proc.time() - ptm
```

**Figure `r iFig=iFig+1; iFig`** Blue whiting: simple projection.

\newpage

```{r}
d=rbind(cbind("What"="M",d1),
        cbind("What"="DD-M",d2),
        cbind("What"="DD-MM",d3),
        cbind("What"="DD-MMM",d4))
ggplot(d)+
  geom_line(aes(SSB,Yield,col=What))
```

# Funding

# References